---
title: "Spam Classification Using a Support Vector Machine in Python and R"
author: "Dillon"
date: "May 2, 2018"
output:
  html_document:
    theme: paper
    number_sections: true
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    #code_folding: show
    #css: styles.css
    #flatly, cosmo, paper
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this project I use a support vector machine (SVM) to classify a preprocessed dataset of emails as spam or not with a 98.9% test set accuracy. A support vector machine is a supervised machine learning model used for classification.


You can download my code at [github](https://github.com/dilloncamp/spam) to follow along. I begin this guide with Python and the scikit-learn package and then perform the same analysis in R with the e1071 package in R.


# Python SVM Walkthrough

## Setup

```{r, echo = FALSE, warning= FALSE, results = 'hide', message=FALSE}
library(shiny)
library(shinydashboard)
```

### Load Packages
```
import scipy.io as sio
from sklearn import svm
import numpy as np
import h5py
import pandas as pd
```

### Load Email Data

```
traindata = sio.loadmat('spamTrain.mat')
X = traindata['X']
y = traindata['y']
##Validation dataset is saved in a different format
##So the process to load it is slightly different
f = h5py.File('spamTest2.mat','r') 
Xtest = np.array(f.get('Xtest'))
Xtest = Xtest.T
ytest = np.array(f.get('ytest'))
ytest = ytest.T
```

The dataset I am using is a preprocessed set of emails from Stanford's Machine Learning course on Coursera. The 1899 X values represent words that appeared at least 100 times in the entire email dataset. For a given email, X will be 1 if that word is present and 0 otherwise. The value Y represents whether the email is spam (Y=1) or not (Y=0).

## Train Support Vector Machine

```
##Train SVM with linear kernel
##Use for loop to determine best value for regularization parameter C
C = [0.01,0.03,0.1,0.3,1,3,10]
trainingAccuracy = np.empty([len(C)])
testingAccuracy = np.empty([len(C)])
f1 = np.empty([len(C)])
for i in range(0,len(C)):
    clf = svm.SVC(C=C[i],kernel='linear')
    clf.fit(X, y.ravel())

    trainPredict = clf.predict(X)
    trainingAccuracy[i] = np.mean(trainPredict==y.ravel())*100
    
    testPredict = clf.predict(Xtest)
    testingAccuracy[i] = np.mean(testPredict==ytest.ravel())*100
    
    #f1 score
    tp = (np.logical_and(testPredict[:] == 1,ytest.ravel()[:] == 1))
    p = ytest.ravel() == 1
    recall = sum(tp)/sum(p)
    fp = (np.logical_and(testPredict[:] == 1,ytest.ravel()[:] == 0))
    precision = sum(tp)/(sum(tp)+sum(fp))
    f1[i] = (2*(precision*recall))/(precision+recall)
    
```

## Results

```
print(testingAccuracy)
Output: array([98. , 99. , 98.9, 98.3, 97.8, 97.3, 97.5])
print(trainingAccuracy)
Output: array([ 98.325,  99.425,  99.825,  99.975,  99.975,  99.975, 100.   ])
print(f1)
Output: array([0.96710526, 0.98381877, 0.98228663, 0.9726248 , 0.96451613, 0.95638126, 0.95974235])
```

Since the test set accuracy is highest for C=.03, I use C=.03 as the regularization parameter. The support vector machine was able to classify the test email set with an accuracy of 99.0% and an F1 score of 98.38%. 


# R Code

## Setup

### Load Packages

```{r, eval = FALSE}
library(e1071)
library(R.matlab)
library(rhdf5)
```

### Load Data

```{r, eval = FALSE}
traindata <- readMat('spamTrain.mat')
X <- traindata$X
Y <- traindata$y
Xtest <- h5read('spamTest2.mat','Xtest')
ytest <- h5read('spamTest2.mat','ytest')
```

## Train Model

```{r, eval = FALSE}
#Using cost determined in Python
clf <- svm(cost = .03, x=X,y=Y,kernel="linear",scale=FALSE,type="C-classification")
trainPredict <- fitted(clf)
trainingAccuracy <- mean(trainPredict == Y)
pred <- predict(clf, Xtest)
testingAccuracy <- mean(pred == ytest)

#f1 score
tp = pred== 1 & ytest == 1
p = ytest == 1
recall = sum(tp)/sum(p)
fp = pred == 1 & ytest == 0
precision = sum(tp)/(sum(tp)+sum(fp))
f1 = (2*(precision*recall))/(precision+recall)
```

## Results

```{r, message = FALSE, echo = FALSE}
load("spamresults.RData")
```

```{r, echo = TRUE}
print(testingAccuracy)
print(f1)
```



